Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem

### 2015-12-10 acta_manuscript.tex
made second set of corrections from dr. lee. regenerated some of the plots

To address the issues associated with conventional numerical integration, we apply computational geometric optimal control techniques. 
We generalize the previous work of determining orbital transfers using invariant manifolds~\cite{koon2011}.
We incorporate the effects of continuous low thrust propulsion to enable departure from the system dynamics and enable increased transfer opportunities.
The dynamics of the three-body system are derived from the discrete Lagrangian, which approximates the integral of the continuous time Lagrangian over a fixed discrete step.
Application of the discrete Euler-Lagrange equations, or equivalently the discrete Legendre transform, results in the discrete equations of motion.
This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs.
A discrete optimal control problem is formulated from the discrete equations of motion.
This approach, where explicit discretization occurs prior to optimization,  is in contrast to the typical method, where the equations of motion are implicitly discretized during the optimization procedure.
Formulating the problem in this manner results in more stable and accurate optimal solutions. 
In indirect methods the optimal control problem is expressed as a two-point boundary value problem.
Optimal solutions are generally sensitive to small variations in the initial multipliers.
As a result, the numerical stability of sensitivity derivatives is critical to accuracy and computational performance. 
The use of geometric integrators, which do not suffer the numerical dissipation of conventional integration methods, results in a more robust and efficient solution.

@@

\subsection{Jacobi Integral}\label{sec:jacobi}
There exists a single integral, or constant of motion for the three-body problem~\cite{lanczos1970,szebehely1967}.
This energy constant is analogous to the total mechanical energy, however it is a non-physical quantity arising from the problem formulation~\cite{szebehely1967}.
Also known as the Jacobi constant, it is defined as a function of the position and velocity in the rotating frame and given by
\begin{equation}
	E\left( \vec{r} , \vec{v} \right) = \frac{1}{2}\left( \dot{x}^2 + \dot{y}^2\right) - U\left(x,y \right) \, .
	\label{eq:jacobi}
\end{equation}
\Cref{eq:jacobi} divides the phase space into distinct regions of allowable motion based on the energy level of the spacecraft.
Fixing the Jacobi integral to a constant defines zero velocity curves, which are the locus of points where the kinetic energy, and hence velocity vanishes.
As seen in Figure~\ref{fig:energy_contour}, the phase space is divided into distinct realms based on the energy level.
In the vicinity of \( m_1\) or \(m_2\) there exists a potential well. 
As the energy level increases there are five critical points of the effective potential~\cref{eq:eff_pot} where the slope is zero.
Three collinear saddle points on the \(\hat{e}_1\) axis and two equilateral points.
These equilibrium, or Lagrange points, are labeled \( L_i, i = 1, \hdots, 5 \) and are shown in Figure~\ref{fig:energy_contour}.
The Jacobi integral is a valuable invariant property of the three-body system that allows for greater insight into the motion of the spacecraft.
\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.5\textwidth]{energy_contours}
	\caption{Contour plot of Jacobi integral}
	\label{fig:energy_contour}
\end{figure} 
%\subsection{Periodic Orbits and Invariant Manifolds}
%Used to generate a Periodic orbit about a lagrange point
%Periodic orbits about the lagrange points
%\subsection{\Poincare Section}
%Intersection of states with a lower dimensional plane
%Introduce invariant manifolds

@@

We define the \Poincare section along the horizontal axis, which is equivalent to the surface \( y(N) = 0 \), and given by
\begin{align}
	\Sigma = \braces{ ( x, \dot{x}) \, | \, y(N) = 0} . 
	\label{eqn:poincare_section}
\end{align}
This is similar to the previous work in determining homoclinic orbits in the three-body problem~\cite{llibre1985,koon2011}.
Previous analytical results have shown that homoclinic orbits intersect transversally in the \( (x, \dot{x} ) \) space on the plane \( y = 0 \).
We seek to compliment these results with the addition of low thrust propulsion to maximize the reachable set on the \Poincare section.
Placing our section at \( y = 0\) ensures that all trajectories will intersect our section.
An optimal control problem is defined by a  cost function
\begin{equation}
	J = -\frac{1}{2} \left( \vec{x}(N) - \vec{x}_{n}(N)\right)^T 
	\left[
	\begin{array}{cccc}
		1 & 0& 0& 0 \\
		 0& 0& 0& 0\\
		 0 & 0 & 1 &0\\
		 0 & 0& 0& 0
	\end{array}
	\right]
	\left( \vec{x}(N) - \vec{x}_{n}(N)\right) \, .
	\label{eq:cost}
\end{equation}
The term \( \vec{x}_n(N) \) is the final state of a control-free trajectory while the term \( \vec{x}(N)\) is the final state under the influence of the control input.
Maximization of the distance between \( \vec{x}_n \) and \(\vec{x} \), in the \Poincare section defined in~\cref{eqn:poincare_section}, is equivalent to the minimization of \( J \).
The \Poincare section is defined through the use of appropriate terminal constraints given by
\begin{subequations}
\begin{align}
    m_1( \vec{x}(N) ) &= 0 = y(N) \, , \\ 
    m_2 ( \vec{x}(N) )&= 0 = \frac{\dot{x}(N) - \dot{x}_n(N) }{x(N) -x_n(N) } - \tan{\theta_d} \, , \\
	 0 &\geq\vec{u}^T \vec{u} - u_{max}^2 \, ,
\end{align}
    \label{eq:constraints}
\end{subequations}
where the angle \( \theta_d\) defines a direction in which we wish to maximize the reachability set on the \Poincare section.
The maximum control thrust magnitude is defined by \( u_{max} \) and is non-dimensionalized by the characteristic units of length, mass, and time.
The goal is to determine the control input \( \vec{u}_k\) such that the cost function~\cref{eq:cost} is minimized subject to the state equations of motion~\cref{eq:discrete_eoms} and constraints~\cref{eq:constraints}.

@@

The optimal control formulation presented in this section results in a two point boundary value problem (TPBVP). 
There exist many methods to solve TPBVPs such as gradient, quasilinearization, and shooting methods~\cite{bryson1975,kirk2012}.
Shooting methods are common in astrodynamic trajectory design problems and relatively simple to implement.
In the shooting method, initial conditions are varied such that a terminal constraint is satisfied, similar to the way an archer modifies the bow in order to more accurately strike a target. 
Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs.
The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches.
Additional interior constraints are used to ensure a continuous solution.
This has the benefit of decreasing the numerical sensitivity of the final states to changes in the initial conditions along each sub-arc.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
The simulations demonstrate the ability of using the reachable set to generate orbital transfer trajectories.
\subsection{Periodic Orbit transfer}\label{sec:periodic_orbit_transfer}
% define problem setup (initial and target orbits)
The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix}^T \) with \( \mu = 0.0125 \).
The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
   \caption{Target Orbit Region}
   \label{fig:moon_orbit}
\end{figure}
\Cref{fig:moon_orbit} shows that the target set remain in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
This type of orbit would be useful for a variety of mission scenarios.
For example, a series of communication satellites could be placed in this type of orbit. 
The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
The initial set is a planar periodic orbit about \( L_1\).
We generate the periodic orbit by using a linear approximation to first determine periodic solutions.
This approximate periodic solution from the linear system is then implemented in the nonlinear dynamics through the method of differential correction, or single shooting~\cite{koon2011}.
The initial \( \dot{y} \) is varied until the the periodic orbit makes a perpendicular crossing, \( \dot{x} = 0 \), after one half period.
Due to the symmetry of the three-body problem, this ensures that the trajectory will return to the initial state after a single period.

@@

% discuss reachability set computation
The optimal trajectories, under the influence of the control input \( \vec{u} \), are plotted in red in~\cref{fig:reach_trajectory}.
Initially, the spacecraft is assumed to lie on the periodic orbit.
As a result, the intersection of this periodic orbit with the \Poincare section are two points corresponding to the two crossing of the orbit.
The use of the continuous low thrust propulsion expands the reachable set to region bounded by the red markers in~\cref{fig:poincare_compare}.
The reachable set is an ellipsoidal region with a major axis aligned along \( \theta \approx \ang{70} \) as compared to a fixed point without any control input.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{reach_trajectory} 
		\caption{Reachability set trajectories} \label{fig:reach_trajectory} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare_compare} 
		\caption{Poincar\`e Section} \label{fig:poincare_compare} 
	\end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare_compare} 
		\caption{Transfer trajectory} \label{fig:reach_transfer} 
	\end{subfigure}~ 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{control_input_l1} 
		\caption{Control input} \label{fig:control_l1} 
	\end{subfigure}~
	\caption{Reachability set transfer}
	\label{fig:reachability_set_transfer} 
\end{figure}

@@

% describe initial and target states for this example
This numerical simulation demonstrates the ability of using multiple iterations of the reachability set to achieve a more complicated transfer.
In this manner, it is possible to design arbitrary transfers which are not possible using a single reachability computation.
Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional Earth-Moon three-body system. 
It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
The initial and target orbits are illustrated in~\cref{fig:geo_transfer_target}.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
   \caption{Transfer target}
   \label{fig:geo_transfer_target}
\end{figure}
In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
We seek to generate a transfer from the geostationary orbit onto the stable invariant manfiold.
Once on the manifold, the spacecraft will coast in an uncontrolled fashion and asymptotically arrive at the desired periodic orbit.

\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{third_reachability_set} % requires the graphicx package
   \caption{Third reachability set}
   \label{fig:third_reachability_set}
\end{figure}
We first generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_{max} = 0.75 \).
The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eq:constraints} defined on the \Poincare section.
This allows us to approximate the set of states that are achievable in the \( \parenth{x ,\dot{x}} \) space.
After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds.
%A state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
For example,~\cref{fig:third_reachability_set} shows the reachable set after three iterations in red and the minimum distance to the stable manifold in green.

With each reachable set we move the controlled trajectory closer to the stable manifold.
\Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 
Combining these trajectories results in the powered portion of the transfer from the geostationary orbit to the stable manifold. 
\Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory, with the final trajectory shown in red which ensures the intersection with the stable invariant manifold.
Once at the stable manifold, no further control input is required and the vehicle will coast towards the target periodic orbit.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_full} 
		\caption{Transfer} \label{fig:geo_transfer_full} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_zoom} 
		\caption{State Trajectory} \label{fig:geo_transfer_zoom} 
	\end{subfigure}

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare} 
		\caption{\Poincare section} \label{fig:geo_transfer_poincare} 
	\end{subfigure}~%
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{control_input_geo} 
		\caption{Control input} \label{fig:control_input_geo} 
	\end{subfigure} 
	\caption{Geostationary Transfer}
	\label{fig:geo_transfer} 
\end{figure}

>>>>

In order to address the issues associated with conventional numerical integration, we apply computational geometric optimal control techniques. 
We generalize the previous work of determining orbital transfers using invariant manifolds~\cite{koon2011}.
We incorporate the effects of continuous low thrust propulsion to enable departure from the system dynamics and enable increased transfer opportunities.
The dynamics of the three-body system are derived from the discrete Lagrangian, which approximates the integral of the continuous time Lagrangian over a fixed discrete step.
Application of the discrete Euler-Lagrange equations, or equivalently the discrete Legendre transform, results in the discrete equations of motion.
This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs.
A discrete optimal control problem is formulated from the discrete equations of motion.
This approach, where explicit discretization occurs prior to optimization,  is in contrast to the typical method, where the equations of motion are implicitly discretized during the optimization procedure.
Formulating the problem in this manner results in more stable and accurate optimal solutions. 
In indirect methods the optimal control problem is expressed as a two-point boundary value problem.
Optimal solutions are generally sensitive to small variations in the initial multipliers.
As a result, the numerical stability of sensitivity derivatives is critical to accuracy and computational performance. 
The use of geometric integrators, which do not suffer the numerical dissipation of conventional integration methods, results in a more robust and efficient solution.

@@

\subsection{Jacobi Integral}\label{sec:jacobi}
There exists a single integral, or constant of motion for the three-body problem~\cite{szebehely1967,lanczos1970}.
This energy constant is analogous to the total mechanical energy, however it is a non-physical quantity arising from the problem formulation~\cite{szebehely1967}.
Also known as the Jacobi constant, it is defined as a function of the position and velocity in the rotating frame and given by
\begin{equation}
	E\left( \vec{r} , \vec{v} \right) = \frac{1}{2}\left( \dot{x}^2 + \dot{y}^2\right) - U\left(x,y \right) \, .
	\label{eq:jacobi}
\end{equation}
\Cref{eq:jacobi} divides the phase space into distinct regions of allowable motion based on the energy level of the spacecraft.
Fixing the Jacobi integral to a constant defines zero velocity curves, which are the locus of points where the kinetic energy, and hence velocity vanishes.
As seen in Figure~\ref{fig:energy_contour}, the phase space is divided into distinct realms based on the energy level.
In the vicinity of \( m_1\) or \(m_2\) there exists a potential well. 
As the energy level increases there are five critical points of the effective potential~\cref{eq:eff_pot} where the slope is zero.
Three collinear saddle points on the \(\hat{e}_1\) axis and two equilateral points.
These equilibrium, or Lagrange points, are labeled \( L_i, i = 1, \hdots, 5 \) and are shown in Figure~\ref{fig:energy_contour}.
The Jacobi integral is a valuable invariant property of the three-body system that allows for greater insight into the motion of the spacecraft.
\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.5\textwidth]{energy_contours}
	\caption{Contour plot of Jacobi integral}
	\label{fig:energy_contour}
\end{figure} 
%\subsection{Periodic Orbits and Invariant Manifolds}
%Used to generate a Periodic orbit about a lagrange point
%Periodic orbits about the lagrange points
%\subsection{\Poincare Section}
%Intersection of states with a lower dimensional plane
%Introduce invariant manifolds

@@

We define the \Poincare section along the horizontal axis, which is equivalent to the surface \( y(N) = 0 \), and given by
\begin{align}
	\Sigma = \braces{ ( x, \dot{x}) \, | \, y(N) = 0} . 
	\label{eqn:poincare_section}
\end{align}
This is similar to the previous work in determining homoclinic orbits in the three-body problem~\cite{llibre1985,koon2011}.
Previous analytical results have shown that homoclinic orbits intersect transversally in the \( (x, \dot{x} ) \) space on the plane \( y = 0 \).
We seek to compliment these results with the addition of low thrust propulsion to maximize the reachable set on the \Poincare section.
Placing our section at \( y = 0\) ensures that all trajectories will intersect our section.
An optimal control problem is defined by a  cost function
\begin{equation}
	J = -\frac{1}{2} \left( \vec{x}(N) - \vec{x}_{n}(N)\right)^T 
	\left[
	\begin{array}{cccc}
		1 & 0& 0& 0 \\
		 0& 0& 0& 0\\
		 0 & 0 & 1 &0\\
		 0 & 0& 0& 0
	\end{array}
	\right]
	\left( \vec{x}(N) - \vec{x}_{n}(N)\right) \, .
	\label{eq:cost}
\end{equation}
The term \( \vec{x}_n(N) \) is the final state of a control-free trajectory while the term \( \vec{x}(N)\) is the final state under the influence of the control input.
Maximization of the distance between \( \vec{x}_n \) and \(\vec{x} \), on the \Poincare section defined in~\cref{eqn:poincare_section}, is equivalent to the minimization of \( J \) defined in~\cref{eq:cost}.
The \Poincare section is defined through the use of appropriate terminal constraints given by
\begin{subequations}
\begin{align}
    m_1( \vec{x}(N) ) &= y(N) = 0 \, , \\ 
    m_2 ( \vec{x}(N) )&=  \frac{\dot{x}(N) - \dot{x}_n(N) }{x(N) -x_n(N) } - \tan{\theta_d} = 0\, , \\
	 0 &\geq\vec{u}^T \vec{u} - u_{max}^2 \, ,
\end{align}
    \label{eq:constraints}
\end{subequations}
where the angle \( \theta_d\) defines a direction in which we wish to maximize the reachability set on the \Poincare section.
The maximum control thrust magnitude is defined by \( u_{max} \) and is non-dimensionalized by the characteristic units of length, mass, and time.
The goal is to determine the control input \( \vec{u}_k\) such that the cost function~\cref{eq:cost} is minimized subject to the state equations of motion~\cref{eq:discrete_eoms} and constraints~\cref{eq:constraints}.

@@

The optimal control formulation presented in this section results in a two point boundary value problem (TPBVP). 
There exist many methods to solve TPBVPs such as gradient, quasilinearization, and shooting methods~\cite{bryson1975,kirk2012}.
Shooting methods are common in astrodynamic trajectory design problems and relatively simple to implement.
In the shooting method, initial conditions are varied such that a terminal constraint is satisfied, similar to the way an archer modifies the bow in order to more accurately strike a target. 
Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs~\cite{stoer2013}.
The multiple shooting method is implemented in this paper which alleviates many of the issues associated with single shooting approaches.
Additional interior constraints are used to ensure a continuous solution.
This has the benefit of decreasing the numerical sensitivity of the final states to changes in the initial conditions along each sub-arc.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
These simulations enable the spacecraft to depart from the natural dynamics through the use of low-thrust propulsion.
The reachability set on the \Poincare section allows for a straightforward method of determining transfer opportunities.

\subsection{Periodic Orbit transfer}\label{sec:periodic_orbit_transfer}
% define problem setup (initial and target orbits)
The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix}^T \) with \( \mu = 0.0125 \).
The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
   \caption{Target Orbit Region}
   \label{fig:moon_orbit}
\end{figure}
\Cref{fig:moon_orbit} shows that the target set remain in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
This type of orbit would be useful for a variety of mission scenarios.
For example, a series of communication satellites could be placed in this type of orbit. 
The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
The initial set is a planar periodic orbit about \( L_1\).
We generate the periodic orbit by using a linear approximation to first determine periodic solutions.
This approximate periodic solution from the linear system is then implemented in the nonlinear dynamics through the method of differential correction, or single shooting~\cite{koon2011}.
The initial \( \dot{y} \) is varied until the the periodic orbit makes a perpendicular crossing, \( \dot{x} = 0 \), after one half period.
Due to the symmetry of the three-body problem, this ensures that the trajectory will return to the initial state after a single period.

@@

% discuss reachability set computation
The optimal trajectories, under the influence of the control input \( \vec{u} \), are plotted in red in~\cref{fig:reach_trajectory}.
Initially, the spacecraft is assumed to lie on the periodic orbit.
As a result, the intersection of this periodic orbit with the \Poincare section are two points corresponding to the two crossing of the orbit.
The use of the continuous low thrust propulsion expands the reachable set to region bounded by the red markers in~\cref{fig:poincare_compare}.
The reachable set is an ellipsoidal region with a major axis aligned along \( \theta \approx \ang{70} \) as compared to a fixed point without any control input.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{reach_trajectory} 
		\caption{Reachability set trajectories} \label{fig:reach_trajectory} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare_compare} 
		\caption{\Poincare section} \label{fig:poincare_compare} 
	\end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{reach_transfer} 
		\caption{Final transfer trajectory} \label{fig:reach_transfer} 
	\end{subfigure}~ 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{control_input_l1} 
		\caption{Control input} \label{fig:control_l1} 
	\end{subfigure}~
	\caption{Reachability set transfer}
	\label{fig:reachability_set_transfer} 
\end{figure}

@@

% describe initial and target states for this example
This numerical simulation demonstrates the ability of the proposed approach. 
We use multiple iterations of the reachability set to achieve a more complicated transfer.
In this manner, it is possible to design arbitrary transfers which are not possible using a single reachability computation.
Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional Earth-Moon three-body system. 
It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
The initial and target orbits are illustrated in~\cref{fig:geo_transfer_target}.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
   \caption{Transfer target}
   \label{fig:geo_transfer_target}
\end{figure}
In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
We seek to generate a transfer from the geostationary orbit onto the stable invariant manfiold.
Once on the manifold, the spacecraft will coast in an uncontrolled fashion and asymptotically arrive at the desired periodic orbit.

\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{third_reachability_set} % requires the graphicx package
   \caption{Reachable set after three iterations}
   \label{fig:third_reachability_set}
\end{figure}
We first generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in~\cref{fig:geo_transfer_full}.
The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_{max} = 0.75 \).
The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eq:constraints} defined on the \Poincare section.
This allows us to approximate the set of states that are achievable in the \( \parenth{x ,\dot{x}} \) space.
After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds.
%A state is chosen from the reachable set which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, and is used as the initial condition for the next reachable set computation.
For example,~\cref{fig:third_reachability_set} shows the reachable set after three iterations in red and the minimum distance to the stable manifold in green.

With each reachable set we move the controlled trajectory closer to the stable manifold.
\Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 
Combining these trajectories results in the powered portion of the transfer from the geostationary orbit to the stable manifold. 
\Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory, with the final trajectory shown in red which ensures the intersection with the stable invariant manifold.
Once at the stable manifold, no further control input is required and the vehicle will coast towards the target periodic orbit.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_full} 
		\caption{Completed transfer trajectory} \label{fig:geo_transfer_full} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_zoom} 
		\caption{Detailed view of transfer} \label{fig:geo_transfer_zoom} 
	\end{subfigure}

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare} 
		\caption{\Poincare section} \label{fig:geo_transfer_poincare} 
	\end{subfigure}~%
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{control_input_geo} 
		\caption{Control input} \label{fig:control_input_geo} 
	\end{subfigure} 
	\caption{Transfer from geostationary orbit to the \( L_1 \) stable manifold using multiple iterations of the reachability set}
	\label{fig:geo_transfer} 
\end{figure}
